\\large Taking uncertainty in word embedding bias estimation seriously \\- a  Bayesian

### 2021-08-02 presentationESSLLI21/presentationESSLLI.Rmd
Modifies presentation
Working on the presentation

---
title: "\\large Taking uncertainty in word embedding bias estimation seriously \\ a  Bayesian
  approach"
author: "Alicja Dobrzeniecka \\& Rafal  Urbaniak \\footnotesize \\newline (LoPSE research group, University of  Gdansk)" 
date:  "ExpSem2021, ESSLLI"
fontsize: 10pt
output:  
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
classoption: x11names, dvipsnames, bibspacing,natbib
urlcolor: blue
bibliography: [../references/cosineReferences1.bib]
csl: ../references/apa-6th-edition.csl 
link-citations: yes
---

@@

- Representation of words (or words in contexts) with vectors of real numbers

@@

@@

@@

### Visual example

@@

##  Bayesian estimation

@@

##  Bayesian estimation

@@

## Effects of debiasing

### Further work

- downstream tasks

<!-- \frametitle{The gist of the project} -->

<!-- \pause -->

<!-- ### Classical probabilism in epistemology (and in applications) -->

<!-- Represent uncertainty by a single probabilistic measure (point estimates). -->

<!-- \vspace{-3mm} -->

<!-- \begin{align*} -->
<!-- \pr(\mathsf{recovery}\vert \mathsf{treatment}) & = .6\\ -->
<!-- \end{align*} -->

<!-- \vspace{-6mm} -->

<!-- \pause -->

<!-- ### Imprecise probabilism -->

<!-- Evidence (e.g. expert evidence) sometimes fails to warrant such precision. -->

<!-- \vspace{-3mm} -->

<!-- \begin{align*} -->
<!-- \pr(\mathsf{recovery}\vert \mathsf{treatment}) & \in [.4,.6]\\ -->
<!-- \end{align*} -->

<!-- \vspace{-6mm} -->

<!-- \pause  -->

<!-- ### Epistemological difficulties of imprecise probabilism -->

<!-- - hard to model the weight of and combine evidence. -->

<!-- - inertia of agnosticism. -->

<!-- - growth of uncertainty  in light of new evidence. -->

<!-- ## The gist of the project -->

<!-- ### My novel contribution -->

<!-- - Meet the desiderata motivating imprecision. -->

<!-- - Avoid these difficulties. -->

<!-- - Do so by using higher-order probabilities.  -->

<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%"} -->
<!-- p <- ggplot(data = data.frame(probability = seq(0,1, by = .25)),mapping = aes(x = probability))+th +ylab("density")+scale_x_continuous(breaks = seq(0,1, by = .1)) -->

<!-- agentPrior <- function(p) dbeta(p,28,28)  -->

<!-- priorPlot <- p + stat_function(fun=agentPrior)+ggtitle("Uncertainty about recovery")+xlab("probability of recovery if treated") -->

<!-- priorPlot -->
<!-- ``` -->

<!-- ## Motivations for imprecision  \large (straighforward considerations) -->

<!-- \pause -->

<!-- ### Radical lack of information -->

<!-- $\pr(\mathsf{It \,\, will \,\,  snow\,\,  in\,\,  Boston\,\,  on \,\, New\,\,  Year's\,\,  Eve\,\,  2026}) =$ ? -->

<!-- \pause -->

<!-- <!-- ### Mere bounds in reasoning -->

<!-- <!-- - Say $\pr(A) = .5$, $\pr(B) = .3$, but you don't know if they're independent. -->

<!-- <!-- \vspace{-2mm} -->

<!-- <!-- - We can only calculate bounds for $\pr(A \et B)$, $\pr(A \vee B)$, etc. -->

<!-- <!-- \pause  -->

<!-- ### Weight of evidence -->

<!-- - You've seen 4/10 patients recover. $\pr(\mathsf{recovery})=.4$? -->

<!-- \pause -->
<!-- \vspace{-2mm} -->

<!-- - You've seen 40/100 patients recover. $\pr(\mathsf{recovery})=.4$? -->

<!-- \footnotesize (You could try to use classical statistical tools, such as confidence intervals,\linebreak  but they have their own problems.) -->

<!-- ## Motivations for imprecision \large (uniformity and agnosticism) -->

<!-- ### Agnosticism preservation requirement -->

<!--  Agnosticism about uniquely connected variables should be the same.   -->

<!-- \vspace{-5mm} -->

<!--  \[X \mbox{ and } Y=X+2 \mbox{ or } Z = X^2 \,\,\, \mbox{ for $10>X>0$.}\]  -->

<!-- \pause -->

<!-- ### The trouble -->

<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%"} -->
<!-- p <- ggplot(data = data.frame(value = 0:10),  -->
<!--             mapping = aes(x = value))  -->

<!-- unifX <-  function(value) punif(value,min = 0, max = 10)  -->

<!-- p +   stat_function(fun=unifX)+ -->
<!--   ylab("density")+theme_tufte(base_size = 25)+labs(title = "Uniform cumulative probability for X in (0,10)")+xlab("X")+ylab("Cumulative probability")+scale_x_continuous(breaks = 0:10)+ -->
<!--   geom_segment(aes(x = 5, y = 0, xend = 5, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")+ -->
<!--   geom_segment(aes(x = 0, y = .5, xend = 5, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue") -->

<!-- ``` -->

<!-- \vspace{-9mm} -->

<!-- \begin{align*}\pr(X<5) = .5 = \pr(X^2 < 25) -->
<!-- \end{align*} -->

<!-- ## Motivations for imprecision \large (uniformity and agnosticism) -->

<!-- ### The trouble -->

<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", message = FALSE, warning = FALSE} -->
<!-- p <- ggplot(data = data.frame(value = 0:10),  -->
<!--             mapping = aes(x = value))  -->

<!-- unifY <-  function(value) punif(sqrt(value),min = 0, max = 10)  -->

<!-- p +   stat_function(fun=unifY)+ -->
<!--   ylab("cumulative probability")+theme_tufte(base_size = 25)+ -->
<!--   labs(title = TeX("Resulting cumulative probability for $X^2$"))+scale_x_continuous(breaks = 0:100)+xlab(TeX("X^2"))+xlim(c(0,100))+ -->
<!--   geom_segment(aes(x = 25, y = 0, xend = 25, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")+ -->
<!--   geom_segment(aes(x = 0, y = .5, xend = 25, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue") -->

<!-- ``` -->

<!-- ## Motivations for imprecision \large (expert evidence) -->

<!-- ### Weather forecasters -->

<!-- \vspace{-1mm} -->

<!-- \begin{tabular}{ll} -->
<!-- Al: & $\pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow})=.6$\\ -->
<!-- Bert: & $\pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow})=.8$ -->
<!-- \end{tabular} -->

<!-- \pause -->

<!-- ### Linear pooling -->

<!-- \vspace{-4mm} -->

<!-- \begin{align*} -->
<!-- \pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow}) & =  -->
<!-- \alpha \times .6 + \beta \times .8 \,\,\,\,\,  (where \,\, \alpha + \beta = 1.) -->
<!-- \end{align*} -->

<!-- \pause -->

<!-- ### Mathematical impossibilities -->

<!-- - Experts's agreement on independence is not preserved by linear pooling. [@List2011Group] -->

<!-- - You can't at the same time hold the following: [@Gallow2018No] -->

<!-- \vspace{-7mm} -->

<!-- \begin{align} -->
<!-- \pr(A=B) & <1\\ -->
<!-- \pr(r\vert A=a) & = a \\ -->
<!-- \pr(r\vert B=b) & = b \\ -->
<!-- \forall a,b \, \pr(r \vert A =a, B = b) & =  \alpha a + \beta b -->
<!-- \end{align} -->

<!-- ## Imprecise probabilism and the motivations -->

<!-- ### The main claim -->

<!-- \justify Represent uncertainty by  a set of probability measures. -->

<!-- \vspace{1mm} -->
<!-- \scriptsize  -->

<!-- [@Levi1974;@Grdenfors1982;@Jeffrey1983-JEFBWA-2;@walley1991statistical;@Kaplan1996;@Joyce2005;@Grove2009;@elkin2017imprecise] -->

<!-- \pause -->

<!-- ### Snowing in Boston?  -->

<!-- \vspace{-1mm} -->

<!-- Say, (.5,1). -->

<!-- \vspace{1mm} -->
<!-- \scriptsize  -->

<!-- [@Mahtani2019Imprecise] -->

<!-- \pause -->

<!-- ### Weight of evidence?  -->

<!-- \vspace{-1mm} -->

<!-- The more evidence you have, the more narrow the interval. -->

<!-- \vspace{1mm} -->
<!-- \scriptsize  -->

<!-- [@Kaplan1996;@Joyce2005;@Sturgeon2008-STURAT] -->

<!-- \pause -->

<!-- <!-- ## Imprecise probabilism and the motivations -->

<!-- ### Agnosticisim? -->

<!-- \vspace{-1mm} -->

<!-- [0,1] interval instead of a single uniform distribution. -->

<!-- \pause -->

<!-- ### Multiple sources?  -->

<!-- \vspace{-1mm} -->

<!-- Just put all those probabilities in one set. -->

<!-- ## Epistemological challenges -->

<!-- \pause -->

<!-- ### Radical uncertainty and weight -->

<!-- Say one experiment observed 40/100  and  another 8/15 recoveries.  -->

<!-- Simply putting the frequencies in a set won't do justice to the evidence.  -->

<!-- \scriptsize  -->
<!-- \vspace{1mm} -->

<!-- [@Bradley2019Imprecise] -->

<!-- \pause -->

<!-- ### Dilation  -->

<!--   \justify Whenever your roomate goes to the track he flips a fair coin. If it comes up heads he bets on  Speedy. Otherwise he places no bets. He comes home from the track grinning.  You're now more uncertain about  $\pr(\mathsf{heads})$ than before.  -->

<!-- \scriptsize  -->

<!-- \vspace{1mm} -->
<!--  [@walley1991statistical;@seidenfeld1993;@White2009-WHIESA;@Topey2012;@pedersen2014demystifying;@Hart2015;@Bradley2016] -->

<!-- ## Epistemological challenges -->

<!-- ### Belief inertia -->

<!-- Say you're fully agnostic about $H$: $P(H) = [0,1]$.  -->

<!-- \begin{center} -->
<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", warning=FALSE, message = FALSE, dpi = 800} -->
<!-- p <- ggplot(data = data.frame(value = 1:7),  -->
<!--             mapping = aes(x = value))+ theme_tufte(base_size = 24)+ -->
<!--   theme(axis.line.x=element_blank(), -->
<!--         axis.text.x=element_blank(), -->
<!--         axis.ticks.x=element_blank(), -->
<!--         axis.title.x=element_blank(), -->
<!--         panel.grid.minor.x=element_blank(), -->
<!--         panel.grid.major.x=element_blank())+ -->
<!--   ylab("probability")+xlim(c(1.8,6.2)) -->

<!-- p + geom_segment(aes(x = 2, y = 0, xend = 2, yend = 1),  -->
<!--                  lineend = "square")+ -->
<!--    geom_segment(aes(x = 6, y = 0, xend = 6, yend = 1),  -->
<!--                  lineend = "square")+ -->
<!--    annotate("text", x = 1.8, y = 0, label = TeX("$p_F$"), size = 7 )+ -->
<!--   annotate("text", x = 1.8, y = 1, label = TeX("$p_T$"), size = 7)+ -->
<!--   annotate("text", x = 1.8, y = .4, label = TeX("$p_{1}$"), size = 7)+ -->
<!--   annotate("text", x = 1.8, y = .6, label = TeX("$p_{2}$"), size = 7)+ -->
<!--   geom_segment(aes(x = 1.95, y = 0, xend = 2.05, yend = 0),  -->
<!--              lineend = "square")+ -->
<!--   geom_segment(aes(x = 1.95, y = 0.4, xend = 2.05, yend = 0.4),  -->
<!--                lineend = "square")+ -->
<!--   geom_segment(aes(x = 1.95, y = 0.6, xend = 2.05, yend = 0.6),  -->
<!--                lineend = "square")+ -->
<!--   geom_segment(aes(x = 1.95, y = 1, xend = 2.05, yend = 1),  -->
<!--                lineend = "square") + #now the second one -->
<!--   annotate("text", x = 6.2, y = 0, label = TeX("$p_F'$"), size = 7)+ -->
<!--   annotate("text", x = 6.2, y = 1, label = TeX("$p_T'$"), size = 7)+ -->
<!--   annotate("text", x = 6.2, y = .2, label = TeX("$p_{1}'$"), size = 7)+ -->
<!--   annotate("text", x = 6.2, y = .5, label = TeX("$p_{2}'$"), size = 7)+ -->
<!--   geom_segment(aes(x = 5.95, y = 0, xend = 6.05, yend = 0),  -->
<!--                lineend = "square")+ -->
<!--   geom_segment(aes(x = 5.95, y = 0.2, xend = 6.05, yend = 0.2),  -->
<!--                lineend = "square")+ -->
<!--   geom_segment(aes(x = 5.95, y = 0.5, xend = 6.05, yend = 0.5),  -->
<!--                lineend = "square")+ -->
<!--   geom_segment(aes(x = 5.95, y = 1, xend = 6.05, yend = 1),  -->
<!--                lineend = "square")+  -->
<!--                  geom_segment(aes(x = 2.1, xend = 5.9, y = 0, -->
<!--                 yend = 0), arrow = arrow(length = unit(0.08, "inches")), -->
<!--                 color = "red", size = 0.6)+  -->
<!--   geom_segment(aes(x = 2.1, xend = 5.9, y = 1, -->
<!--                    yend = 1), arrow = arrow(length = unit(0.08, "inches")), -->
<!--                color = "red", size = 0.6)+  -->
<!--   geom_segment(aes(x = 2.1, xend = 5.9, y = 0.4, -->
<!--                    yend = .2), arrow = arrow(length = unit(0.08, "inches")), -->
<!--                color = "red", size = 0.6)+  -->
<!--   geom_segment(aes(x = 2.1, xend = 5.9, y = 0.6, -->
<!--                    yend = .5), arrow = arrow(length = unit(0.08, "inches")), -->
<!--                color = "red", size = 0.6)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.1, yend = 0.06), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.2, yend = 0.1), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.3, yend = 0.14), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+ -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.45, yend = 0.28), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                                                             color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.5, yend = 0.36), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.55, yend = 0.45), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+ -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.65, yend = 0.55), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.7, yend = 0.65), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.75, yend = 0.72), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.79, yend = 0.8), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.84, yend = 0.86), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3)+  -->
<!--   geom_segment(aes(x = 2.2, xend = 5.8, y = 0.9, yend = 0.93), arrow = arrow(length = unit(0.06, "inches")), -->
<!--                color = "skyblue", size = 0.3) -->
<!-- ``` -->
<!-- \end{center} -->

<!-- <!-- \begin{center} --> 
<!-- <!-- \includegraphics[width = 6cm]{inertia.jpg} --> 
<!-- <!-- \end{center} -->

<!-- Then $\pr(H \vert E) =[0,1]$, no matter what the evidence says.  -->

<!-- \vspace{1mm} -->
<!-- \scriptsize -->

<!-- [@Joyce2010,@Dodd2012;@Rinard2013;@Bradley2014;@Vallinder2017;@Moss2020] -->

<!-- ## Epistemological challenges -->

<!-- ### Mystery coins -->

<!-- - A fair draw of one of two biased coins, with biases $.4, and .6$. -->

<!-- \pause -->

<!-- \vspace{-2mm} -->

<!-- - Imprecise reaction: $\{.4, .6\}$.  -->

<!-- \pause -->

<!-- \vspace{-2mm} -->

<!-- - Precise probabilism: expected bias of $.5$. -->

<!-- \pause -->

<!-- ### Question -->

<!-- As far as accuracy is concerned, why prefer the imprecise way? -->

<!-- Say the actual bias is .4. A first stab: -->

<!-- \vspace{-3mm} -->

<!-- \begin{align*} -->
<!-- \mathsf{inaccuracy(imprecise)} & = \frac{0 + .2}{2} = .1\\ -->
<!-- \mathsf{inaccuracy(precise)} & =  .1\\ -->
<!-- \end{align*} -->

<!-- \vspace{-5mm} -->
<!-- \scriptsize -->

<!-- [@Seidenfeld2012forecasting; @MayoWilson2016; @schoenfield2017AccuracyRationalityImprecise] -->

<!-- ## Higher-order approach -->

<!-- ### The mystery coin with  $(.2,.8)$ and 1 head in five tosses -->

<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%"} -->
<!-- theta <- c(.4, .6) #possible parameter values -->
<!-- priorTheta <- c(.2,.8) #priors assigned to them -->
<!-- data <- c(1,0,0,0,0) #1 heads with five tosses -->
<!-- z = sum(data) # number of 1's in Data -->
<!-- N = length(data) #number of tosses -->

<!-- # Compute the Bernoulli likelihood at each value of Theta: -->
<!-- pDataIfTheta <- dbinom(z,N,theta) -->
<!-- # Compute the probability of th evidence and  -->
<!-- # the posterior via Bayes' rule: -->
<!-- pData <- sum(pDataIfTheta * priorTheta) -->
<!-- pThetaIfData <- pDataIfTheta * priorTheta / pData -->

<!-- #put together -->
<!-- table <- data.frame(theta,  priorTheta, -->
<!--                      pData,  pDataIfTheta, pThetaIfData) -->

<!-- #plot of the prior -->
<!-- prior <- ggplot(table, aes(x = theta, y = priorTheta))+   -->
<!--   geom_bar(stat="identity", alpha = .8)+  -->
<!--   scale_x_continuous(breaks = theta)+th+ylim(c(0,1))+ggtitle("Prior") -->

<!-- #plot of the likelihood -->
<!-- likelihood <-  ggplot(table, aes(x = theta, y = pDataIfTheta))+ -->
<!--   geom_bar(stat="identity", alpha = .8)+  -->
<!--   scale_x_continuous(breaks = theta)+ -->
<!--   th+ylim(c(0,1))+ggtitle("Likelihood") -->

<!-- #plot of the posterior -->
<!-- posterior <- ggplot(table, aes(x = theta, y = pThetaIfData))+  #set up -->
<!--   geom_bar(stat="identity", alpha = .8)+  #barplot layer -->
<!--   scale_x_continuous(breaks = theta)+ -->
<!--   th+ylim(c(0,1))+ggtitle("Posterior") -->

<!-- #convert the table to graphics, transpose for readability -->
<!-- tableGG <- ggtexttable(t(round(table,3))) -->

<!-- #plot together -->
<!-- ggarrange(prior, likelihood, posterior, tableGG) -->
<!-- ``` -->

<!-- ## Higher-order approach -->

<!-- ### Two experts, uniform prior -->

<!-- ```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%"} -->
<!-- p <- ggplot(data = data.frame(probability = seq(0,1, by = .25)),mapping = aes(x = probability))+th +ylab("density")+scale_x_continuous(breaks = seq(0,1, by = .25))+ylim(c(0,10)) -->

<!-- agentPrior <- function(p) dbeta(p,1,1)  -->
<!-- al <- function(p) dbeta(p,5,20)  -->
<!-- bert <- function(p) dbeta(p,8,12)  -->
<!-- posterior <- function(p) dbeta(p,19,62)  -->

<!-- priorPlot <- p + stat_function(fun=agentPrior)+ggtitle("Agent\'s prior") -->
<!-- alPlot <- p + stat_function(fun=al)+ggtitle("If Al says .2") -->

<!-- bertPlot <- p + stat_function(fun=bert)+ggtitle("If Bert says .4") -->
<!-- posterior <- p + stat_function(fun=posterior)+ggtitle("Agent's posterior") -->

<!-- ggarrange(priorPlot, alPlot, bertPlot, posterior) -->
<!-- ``` -->

<!-- ## The key conjecture -->

<!-- ### With higher-order methods, it will be possible to: -->

<!-- - match the representation of uncertainty with the evidence, -->

<!-- - combine various sources of uncertainty without running into the known problems, -->

<!-- - use existing mathematics to make sense of accuracy in such contexts, -->

<!-- - capture reasoning with radical uncertainty without dilation and belief inertia. -->

<!-- ## Tasks (discussion) -->

<!-- \footnotesize  -->

<!-- - Relate this representation to weight of evidence. -->

<!-- \vspace{-2mm} -->

<!-- - Go beyond grid approximation and beta representation. -->

<!-- - Represent  agnosticism so that: -->

<!--     - agnosticism  preservation holds, and -->

<!--     - inertia fails.  -->

<!-- \vspace{-2mm} -->

<!-- - Represent opinion aggregation more generally, compare to known limitations and features of already studied opinion pooling methods.  -->

<!-- \vspace{-2mm} -->

<!-- - Investigate dilation-like phenomena in this context.  -->

<!-- \vspace{-2mm} -->

<!-- - Study known distance functions between probabilistic measures as candidates for key elements of accuracy measures. -->

<!-- \vspace{-2mm} -->

<!-- - Explain away the apparent contrast between evidence-grounding and accuracy considerations [@easwaranAndFitelson2015]. -->

<!-- \begin{center} -->
<!-- \Large Thank you for your attention! -->
<!-- \end{center} -->

>>>>

---
title: "\\large Taking uncertainty in word embedding bias estimation seriously \\- a  Bayesian
  approach"
author: "Alicja Dobrzeniecka \\& Rafal  Urbaniak \\footnotesize \\newline (LoPSE research group, University of  Gdansk)" 
date:  "ExpSem2021, ESSLLI"
fontsize: 10pt
output:  
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
classoption: x11names, dvipsnames, bibspacing,natbib
urlcolor: blue
bibliography: [../references/cosineReferences1.bib]
csl: ../references/apa-6th-edition.csl 
link-citations: yes
---

@@

- Representation of words with vectors of real numbers

@@

<!-- check if the statement of co-occurence always right-->

@@

## Cosine-based measures of bias - visual example

@@

@@

- It throws away information about sample sizes
- It removes variation which may result in false confidence

##  Cosine distance and word connection visualization

### Analysis of a word "muslim"

@@

## Cosine distance and word connection visualization

### Analysis of a word "priest"

@@

## Advantages of including uncertainty

- It enables one to directly observe the influence of sample sizes
- It may influence risk assessment and decision making
- ...

## Bayesian model

### Choosing priors

```{r priorsVis,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, out.width = "100%", fig.caption = "Regularizing priors for the bayesian models."}
library(ggpubr)
sim <- seq(-1,1,by = 0.001)
dis <- 1-sim

p <- ggplot(data = data.frame(distance = dis), 
            mapping = aes(x = distance))+theme_tufte()

none <- function(x) dnorm(x,1,.5)
cau <- function(x) dcauchy(x,0,1)
par <- function(x) dnorm(x,0,1)

noneG <- p + stat_function(fun = none)+xlim(c(-0.5,2.5))+ggtitle("Prior for mean distances")
parG <- p + stat_function(fun = par)+xlim(c(-2,2))+xlab("distance change")+ggtitle("Prior for coefficients")
cauG <- p + stat_function(fun = cau)+xlim(c(0,2))+ggtitle("Prior for standard deviation")+xlab("distance change")
ggarrange(noneG,cauG, parG, ncol = 1)
```

##  Dataset-level HPDIs coefficients

### Religion coefficients

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=7cm]{../images/religionCoeffs.jpeg}
   \end{minipage}
  \end{figure}

\end{center}

##  Dataset-level HPDIs coefficients

### Gender coefficients

\begin{center}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/genderCoeffs.jpeg}
\end{minipage}

\end{figure}

\end{center}

## Uncertainty included in bias detection

### Closer look at Gender class

\includegraphics[width=8cm]{../images/visGenderReddit.png}

## Summary of the research

-

## Further work

- including contrasts in Bayesian model
- downstream tasks
- applying uncertainty to WEAT metric
- testing AIT dataset
